---
title: "Computing Dynamic SWATH Windows"
author: "Witold E. Wolski, Bernd Roschitzki"
date: "`r Sys.Date()`"
package: "`r packageVersion('prozor')`"
vignette: >
  %\VignetteIndexEntry{Computing Dynamic SWATH Windows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  html_document
---

__!!!Caution work in progress!!!__

# Introduction

Function optimizes Extraction windows so we have the same number of precursor per window. To do it uses spectral library or non redundant blib.

# Prerequisites

```{r echo=FALSE}
library(prozor)

```

# Constant with method


```{r fig.retina=3}
cdsw <- Cdsw(masses)
cdsw$plot()

```

```{r, results='markup'}
knitr::kable(cdsw$asTable())
```

## Error

```{r}
constError <-cdsw$error()
```

# Classical Method based on `quantile`

Same number of MS1 precursors in each window

```{r}
cdsw$quantile_breaks()
```

```{r fig.retina=3, warning=FALSE}
cdsw$plot()
```

```{r, results='markup'}
knitr::kable(cdsw$asTable())
```

## Error

```{r}
quantileError <- cdsw$error()
```

## Adjust windows

Shifts window start and an to a mass range with few MS1 peaks.

```{r, }
knitr::kable(cdsw$optimizeWindows(maxbin=10,plot=TRUE))
```

# Iterative Distribution Mixing based cdsw

## Requirements

- Mass range can be specified (*mass\_range*)
- Maximal window size can be specified (*max\_window\_size*). This is because windows should not be to large because of optimal collision energy (personal communication by Bernd R.).
- Minimal window size can be specified (*min\_window\_size*).
- target number of windows can be specified (*nr\_windows*).
- boundaries between windows are placed in regions were no precursors are observed.


```{r fig.retina=3}
cdsw$sampling_breaks(maxwindow = 100,plot = TRUE)
```

```{r fig.retina=3}
cdsw$plot()
```

```{r, results='markup'}
knitr::kable(cdsw$asTable())
```

## Error

```{r}
mixedError <-cdsw$error()
```


# Errors

```{r fig.retina=3}
barplot(c(const = constError$score1, quantile = quantileError$score1, mixed = mixedError$score1),ylab="Manhattan distance")
barplot(c(const = constError$score2, quantile = quantileError$score2, mixed = mixedError$score2),ylab="Euclidean distance")

```

# Session info

Here is the output of `sessionInfo()` on the system on which this
document was compiled:

```{r sessionInfo, echo=FALSE}
sessionInfo()
```





