---
title: "Creating Target Decoy databases"
author: "Witek Wolski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Creating Target Decoy databases}
-->


Using the package prozor for creating standarized fasta.
prozor can be used to: 
     
     * merge several fasta files into a single fasta file
     * add reverse sequences to the fasta file.
     * add contaminants to the fasta file


# Contaminants

Here at the FGCZ we have a set of contaminants, and they are also provided with this package.
There are two versions of this contaminant set. One with and one without contaminants of human orgin.
We usually are adding one of those two contaminant files to the database.
The contaminants are easy to distinguish thanks to the `zz|FGCZCont` prefix.

```{r}
library(prozor)
head(names(loadContaminantsFasta()))
length(loadContaminantsFasta())
length(loadContaminantsNoHumanFasta())

```

# Creating a fasta protein amino acid sequence database for searching.


To merge several fasta databases into a single file place them into a single folder and give the folder the name of the database.
The database name consists of the project number e.g. `p1000` a consecutive number e.g. `db1` and descriptive name `example` i.e. `p1000_db1_example`.

Add to the folder also an annotation.txt file. The annotation file should contain a single line formatted like a fasta protein header with the following conent:
`aa|p<project_number>_<database_name>|<YYYYMMDD> <detailed_description>`.

Example : `AA|p1000_db1_example|20180119_Example https://github.com/protViz/prozor`


In the package we provided such an example in the folder "p1000_db1_example"

```{r}


databasedirectory <- file.path(find.package("prozor"), "p1000_db1_example")
dbname <- basename(databasedirectory)
fasta <- grep("fasta", dir(databasedirectory),value = T)
files1 <- file.path(databasedirectory,fasta)
annot <- grep("annotation",dir(databasedirectory), value=T)
annotation <- readLines(file.path(databasedirectory,annot))
annotation
```


## Create non decoy database

```{r}
resDB <- createDecoyDB(files1, useContaminants = loadContaminantsFasta(), annot=annotation, revLab = NULL)
length(resDB)
```

```{r}
dirname(databasedirectory)
xx <- file.path(dirname(databasedirectory), paste(dbname,"_",format(Sys.time(), "%Y%m%d"),".fasta" ,sep=""))
```

```{r eval=FALSE}
writeFasta(resDB, file=xx)
```


## Create decoy database


```{r}
resDBDecoy <- createDecoyDB(files1,useContaminants = loadContaminantsFasta(), annot=annotation, revLab = "_REV")
length(resDBDecoy)

sum(duplicated(names(resDBDecoy)))
sum(duplicated(resDBDecoy))

dbname_decoy <- unlist(strsplit(dbname,"_"))
dbname_decoy <- paste(c(dbname_decoy[1],"d",dbname_decoy[2:length(dbname_decoy)]),collapse="_")
dbname_decoy

xx <- file.path(dirname(databasedirectory), paste(dbname_decoy,"_",format(Sys.time(), "%Y%m%d"),".fasta" ,sep=""))
print(xx)
```


```{r eval=FALSE}

writeFasta(resDBDecoy, file=xx)

```
