---
title: "Target Decoy FDR from COMET search results"
author: "Witold Wolski"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{Target Decoy FDR from COMET search results}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
```

Here we describe how you can estimate the false discovery rate (FDR) when matching your spectra against a FASTA sequence database using spectra search engines such as Mascot or Comet. The sequence database contains forward sequences and the same number of reverse sequences.

## The data

Comet's E-value significantly outperforms SEQUEST'sprob-. ability score and Comet's cross-correlation score performs.


```{r}
library(dplyr)
library(prozor)
library(ggplot2)
#path <- "/Users/witoldwolski/__checkout/HeCaTos/data/autoQC4L/FUSION_1.20180713_005_autoQC4L.HCD.lowres.comet.txt"
#bcd <- read.table(file = path, header= TRUE, skip=1 , fill = TRUE)

data(bcd)
bcd |> ggplot(aes(x = -log10(e.value), y = sp_score)) + geom_point()
bcd |> select(scan , plain_peptide, protein, e.value, sp_score) |> head() |> knitr::kable()

```



# Define helper function

```{r}
summarizeLevel <- function(data,
                           level = "PSM",
                           score = "e.value",
                           protein = "protein",
                           qValue = 1,
                           rev = "REV_"){
    fdr1 <- prozor::computeFDR((data[[score]]), grepl(rev,data[[protein]]),larger_better = FALSE)
    score01K <- prozor::predictScoreFDR(fdr1,qValue = qValue, method = "SimpleFDR")
    
    filt <- dplyr::filter(data, !!sym(score) < score01K)
    res <- data.frame(n = nrow(data),
               nConfident = nrow(filt),
               nDecoy = sum(grepl(rev, filt[[protein]])),
               fdr = mean(grepl(rev, filt[[protein]])),
               assignmentRate = nrow(filt)/nrow(data) * 100)
    colnames(res) <- paste0(colnames(res), level)
    return(res)
}

```

# PSM level

```{r}
summarizeLevel(bcd)

```


# Peptide FDR

Use minimum of e.value of all PSMs matching a peptide as new score for the peptide.

```{r}
peptideLevel <-  bcd |> dplyr::group_by(plain_peptide,protein ) |>
    summarize(n = n(), min_e.value = min(e.value))
```

```{r}
table(peptideLevel$n)
summarizeLevel(peptideLevel, score = "min_e.value", level = "Peptides")

```

# Protein FDR

Use minimum of e.value of all PSMs matching a protein as new score for the protein.

```{r}
proteinLevel <-  bcd |> dplyr::group_by(protein ) |> 
    summarize(n = n(), min_e.value = min(e.value))
```

```{r}
table(proteinLevel$n)
summarizeLevel(proteinLevel, score = "min_e.value",level="Proteins")

```

# Filter on PSM level only

Filter on PSM level and then estimate the FDR on peptide and protein level.

```{r}
filtLevel <- function(data,
                      level = "PSM",
                      score = "e.value",
                      larger_better = FALSE,
                      protein = "protein",
                      qValue = 1,
                      rev = "REV_"){
    fdr1 <- prozor::computeFDR((data[[score]]), grepl(rev,data[[protein]]),larger_better = larger_better)
    score01K <- prozor::predictScoreFDR(fdr1,qValue = qValue, method = "SimpleFDR")
    print(score01K)
    if(larger_better){
        filt <- dplyr::filter(data, !!sym(score) > score01K)
    } else {
        filt <- dplyr::filter(data, !!sym(score) < score01K)
    }
    return(filt)
}


```

## Use e-value

```{r}


# psm level FDR
psmFdrCutoff <- 0.01
#debug(filtLevel)

filt <- filtLevel(bcd, score = "e.value", larger_better = FALSE , qValue = psmFdrCutoff * 100)

res <- list()
res$nPSM <- nrow(bcd)
res$psmFdrCutoff <- psmFdrCutoff
res$fdrPSM <- round(mean(grepl("REV_",filt$protein)) * 100, digits = 2)
res$nConfidentPSM <- nrow(filt)
res$nDecoyPSM <- sum(grepl("REV_",filt$protein))

# peptide level FDR
filtPep <- filt |> select(plain_peptide, protein) |> distinct()
res$fdrPeptide <- round(mean(grepl("REV_",filtPep$protein)) * 100, digits = 2)
res$nConfidentPeptide <- nrow(filtPep)
res$nDecoyPeptide <- sum(grepl("REV_",filtPep$protein))

# protein level FDR
filtProt <- filt |> select( protein) |> distinct()
res$fdrProtein <- round(mean(grepl("REV_",filtProt$protein)) * 100, digits = 2)
res$nConfidentProtein <- nrow(filtProt)
res$nDecoyProtein <- sum(grepl("REV_",filtProt$protein))


knitr::kable(data.frame(res))

```


## Use sp_score

```{r}

# psm level FDR
psmFdrCutoff <- 0.01
filt <- filtLevel(bcd, score = "sp_score", larger_better = TRUE, qValue = psmFdrCutoff * 100)
res <- list()
res$nPSM <- nrow(bcd)
res$psmFdrCutoff <- psmFdrCutoff
res$fdrPSM <- round(mean(grepl("REV_",filt$protein)) * 100, digits = 2)
res$nConfidentPSM <- nrow(filt)
res$nDecoyPSM <- sum(grepl("REV_",filt$protein))

# peptide level FDR
filtPep <- filt |> select(plain_peptide, protein) |> distinct()
res$fdrPeptide <- round(mean(grepl("REV_",filtPep$protein)) * 100, digits = 2)
res$nConfidentPeptide <- nrow(filtPep)
res$nDecoyPeptide <- sum(grepl("REV_",filtPep$protein))

# protein level FDR
filtProt <- filt |> select( protein) |> distinct()
res$fdrProtein <- round(mean(grepl("REV_",filtProt$protein)) * 100, digits = 2)
res$nConfidentProtein <- nrow(filtProt)
res$nDecoyProtein <- sum(grepl("REV_",filtProt$protein))


knitr::kable(data.frame(res))

```


## Using protViz

```{r}
knitr::kable(protViz::summary.cometdecoy(bcd,psmFdrCutoff = 0.01))

```


# DEBUG

```{r}
predictScoreFDR_V2 <- function (fdrObj, qValue = 1, method = "SimpleFDR") 
{
    if (method == "FPR") {
        validFDR <- fdrObj$qValue_FPR < qValue/100
    }
    else if (method == "SimpleFDR") {
        validFDR <- fdrObj$qValue_SimpleFDR < qValue/100
    }
    else {
        stop("no such method: ", method, "\n")
    }
    invalidScores <- fdrObj$score[!validFDR]
    if (!fdrObj$larger_better) {
        min(invalidScores)
    }
    else {
        max(invalidScores)
    }
}
```


